### 2025-04-25 thesis/thesis.tex
Rule based

@@

% Welche approaches fuer llm integration
% Welche approaches fuer natural to sql

>>>>

% bessere einleitung

@@

\subsection{Traditional NL2SQL Approaches}

Prior to the wide-spread dominance of machine learning approaches for natural language processing a variety of
traditional, rather discrtete approaches have been explored in the field of NL2SQL / NLIDBs. These discrete
approaches have laid the foundations for transitioning towards the application of machine learning techniques
for NL2SQL.

\subsubsection{Rule-based and Grammar-based Systems}

Early research of NL2SQL system mostly focused around applying rule engines that were tedious to set up and expensive
to maintain / transfer across database systems. These rule engines mostly relied on the systematic identification
of linguistic patterns / were trying to template SQL from information that was derived from processing the natural
language query. \citep*{Rendevouz, Lunar, Ladder} These approaches mostly tried to formalize natural language queries into 
formal grammers which could then be deterministically mapped into a valid SQL query. These approaches have strong downsides
when it comes to the variety of natural language constructs they can process, aswell as runtime adoption of new / unknown
databases, query constructs etc. A potential upside of this class of NL2SQL systems is that they can confidently and 
reproducably identify questions they can, and can't answer — thus leading to very reliable and predictable user interfaces.

\subsection{Neural NL2SQL Approaches}

\subsection{Benchmark Evolution}

\subsection{LLM-based NL2SQL Systems}

\subsection{Optimization for LLM-based NL2SQL}

\subsection{Complex Techniques in NL2SQL}

\subsection{Research Gaps}
